Meta-analytic approaches and 'new' effect sizes to account for heterogeneous effects in comparative physiology

### 2021-05-18 MS/ms.Rmd
small changes

Meta-analysis is a well-developed research field with a long history in the fields of statistics, medicine, psychology and ecology [@Gurevitch2018; @Nakagawa2017]. It has emerged as the 'gold-standard' across various disciplines for quantitative synthesis, and is becoming increasingly more prominant the field of comparative physiology. Meta-analysis involves strict literature-based search protocols and data extraction [i.e., following the PRISMA protocol, see @ODea2021] to achieve two major goals, 1) to provide an overall mean estimate of a treatment effect or relationship and 2) to quantify effect size variance and understand key drivers explaining differences in effects across studies [@Nakagawa2017; @Gurevitch1999; @Gurevitch2018; @Cooper2009; @Koricheva2013]. A critically important feature of meta-analaysis is its ability to account for sampling error []. The inverse sampling error is used to weight higher powered studies in the analysis and make overall esatimtes more precise. While many meta-analyses might employ unweighted analyses, these studies will not gain greater precision on esatimytes, even if the overall mean itself is unbiased.

@@

Equations \@ref(eq:lnq10VR) and \@ref(eq:slnq10VR) describe the change in physiological rate variance (eqn \@ref(eq:lnq10VR)) across a 10&deg;C temperature change along with its sampling variance (eqn \@ref(eq:slnq10VR)). While this is a useful metric, as discussed by Nakagawa *et al* @nakagawa2015meta there is often a strong mean-variance relationship that needs to be accounted for in analysing changes in variance. As such, we can calculate the coefficient of variation, which standardises changes in variance for changes in means as follows:

@@

It will not possible to derive an effect size that completely controls for known heterogenity at the effect size level all the time. This may because because treatment effects may not be simple linear or exponential functions of the mean response making the relationship with effect size complex [e.g., thermal performance curves; @Noble2018]. For example, dosages applied to treatments can follow quite complex non-linear relationships, with expoential or quadratic relationships in relation to some mean response. When this occurs we recommend comparative physiologists apply meta-regerssion approaches, controlling for these known treatment differences as a moderator. There are simple statisical approaches a meta-analyst can use to then make the overall effects interpretable [@Schielzeth2010]. Before diving into meta-regression models, we will first overview the meta-analytical approaches or steps commonly used.

#### *Multi-level Meta-Analysis when the Overall Effect is of Interest*
One of the main goals of meta-analysis is to obtain an overall meta-analytic mean estimate (and some form of uncertainty around that mean, like a 95% confidence interval, may be even with a prediction interval - @Nakagawa2021c) which describes the magnitude of effect across all the studies / effects included in the meta-analysis. This is generally done with a simple random effects or multilevel meta-analytic model that accounts for effect size sampling error [@Nakagawa2017; @NakagawaSantos2012]. More often than not, meta-analyses in comparative physiology will have complex hierachical structure. For example, multiple effect sizes may be taken from a single study on different traits measured on the same individuals, or come from many different species that share an evolutionary history [@Noble2017]. As such, it is far more likely that comparative physiologists will need to apply a multi-level meta-analytic model to control for the various sources of non-independence [@NakagawaSantos2012; @Song2021; @Nakagawa2021; @Noble2017]. Multilevel models are also extremely useful in describing the various sources contributing to effect size variance. For the purpose of our discussion, we will assume the meta-analyst plans to use more advanced multilevel models. Having said that, it maybe sufficient for many meta-analyses to make use of traditional random effects models. We refer readers to excellent reviews and books that overview the distinctions between common (fixed) effect, random effect and multilevel models and when one might want (or not want) to apply these models [@Nakagawa2017; @NakagawaSantos2012; @Koricheva2013; @Borenstein2009].

>>>>

Meta-analysis is a well-developed research field with a long history in statistics, medicine, psychology and ecology [@Gurevitch2018; @Nakagawa2017]. It has emerged as the 'gold-standard' across various disciplines for quantitative synthesis, and is becoming increasingly more prominant the field of comparative physiology. Formal meta-analyses involve a systematic literature-based search, extraction of data from studies meeting specific inclusion criteria, and the subsequent calculation of effects sizes from summary statistics provided by papers. The entire process should follow and report upon this entire process carefully to ensure the meta-analysis is as transparent and reproducible as possible [i.e., see the Preferred Reporting Items for Systematic Reviews and Meta-Analyses in Ecology and Evolution (PRISMAEcoEvo) protocol, @ODea2021]. Aside from providing an up-to-date overview of the research question, along with identifying any gaps and biases, there are two major goals of meta-analysis: 1) to provide an overall mean estimate of a treatment effect or relationship and 2) to quantify effect size variance and understand key drivers explaining differences in effects across studies [@Nakagawa2017; @Gurevitch1999; @Gurevitch2018; @Cooper2009; @Koricheva2013]. A critically important feature of meta-analaysis is its ability to account for sampling error []. The inverse sampling error is used to weight higher powered studies in the analysis and make overall esatimtes more precise. While many meta-analyses might employ unweighted analyses, these studies will not gain greater precision on esatimytes, even if the overall mean itself is unbiased.

@@

Equations \@ref(eq:lnq10VR) and \@ref(eq:slnq10VR) describe the change in physiological rate variance (eqn \@ref(eq:lnq10VR)) across a 10&deg;C temperature change along with its sampling variance (eqn \@ref(eq:slnq10VR)). While this is a useful metric, as discussed by @Nakagawa2015 there is often a strong mean-variance relationship that needs to be accounted for in analysing changes in variance. As such, we can calculate the coefficient of variation, which standardises changes in variance for changes in means as follows:

@@

It will not be possible to derive an effect size that completely controls for known heterogenity at the effect size level all the time. This may because because treatment effects may not be simple linear or exponential functions of the mean response making the relationship with effect size complex [e.g., thermal performance curves; @Noble2018]. For example, dosages applied to treatments can follow quite complex non-linear relationships, with expoential or quadratic relationships in relation to some mean response. When this occurs we recommend comparative physiologists apply meta-regerssion approaches, controlling for these known treatment differences as a moderator. There are simple statisical approaches a meta-analyst can use to then make the overall effects interpretable [@Schielzeth2010]. Before diving into meta-regression models, we will first overview the meta-analytical approaches or steps commonly used.

#### *Multi-level Meta-Analysis when the Overall Effect is of Interest*
One of the main goals of meta-analysis is to obtain an overall meta-analytic mean estimate (and some form of uncertainty around that mean, like a 95% confidence interval, may be even with a prediction interval - @Nakagawa2021c) which describes the magnitude of effect across all the studies / effects included in the meta-analysis. This is generally done with a simple random effects or multilevel meta-analytic model that accounts for effect size sampling error [@Nakagawa2017; @NakagawaSantos2012]. More often than not, meta-analyses in comparative physiology will have complex hierachical structure. For example, multiple effect sizes may be taken from a single study on different traits measured on the same individuals, or come from many different species that share an evolutionary history [@Noble2017]. As such, it is far more likely that comparative physiologists will need to apply a multi-level meta-analytic model to control for the various sources of non-independence [@NakagawaSantos2012; @Song2021; @Nakagawa2021; @Noble2017]. Multilevel models are also extremely useful in describing the various sources contributing to effect size variance. For the purpose of our discussion, we will assume the meta-analyst plans to use more advanced multilevel models. Having said that, it maybe sufficient for many meta-analyses to make use of traditional random effects models. We refer readers to excellent reviews and books that overview the distinctions between common (fixed) effect, random effect and multilevel models and when one might want (or not want) to apply these models [@Nakagawa2017; @NakagawaSantos2012; @Koricheva2013; @Borenstein2009; @Schmid2021].
